Distributional Regression Forests for Probabilistic Precipitation Forecasting in Complex Terrain

### 2018-10-11 pkg/disttree/inst/draft/draft.Rnw
some more revision based on reviews
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/partykit@2322 255f4249-2263-4417-b1ea-8498779e27a7

Former-commit-id: 0dfc728e986b921f0e6d64e66120d4f400f19891

To assess the predictive performance, a full cross-validation is carried out
rather than relying on just the one fixed test set for the years 2009--2012.
\hl{To do so, a 10 times 7-fold cross-validation is carried out where each
splits the available 28~years into 7~subsets of 4~randomly selected and 
therefore mostly not subsequent years.}
The models are learned on 6~folds (= 24~years) and evaluated on the 7-th fold
(= 4~years) using the average CRPS across all observations.
The resulting 10~CRPS skill scores are displayed by boxplots in
Figure~\ref{fig:axams-crps} using EMOS as the reference model (horizontal
line at a CRPSS of 0). Both GAMLSS models and the distributional forest 
perform distinctly better than the EMOS model. While the two GAMLSS lead to an improvement 
of around 4~percent, the distributional forest has a slightly higher improvement
of around 5.5~percent in median.

@@

\subsubsection*{Variable Selection}
\hl{Generally, there are many possibilities how to specify the variables that are
to be included in a model. Especially for a low number of variables all types
of interactions can be employed and mixed effect models and separations into 
main effects and interaction effects can be considered allowing for easier 
interpretations} (see for example \cite{Wood+Scheipl+Faraway:2013}, 
\cite{Goicoa+Adin+Ugarte:2018}, \cite{Ugarte+Adin+Goicoa:2017}).
However, in the application presented in Section~\ref{sec:precipitation}
alltogether 80 variables are available. Even if only paired interactions were
considered this would already lead to 3160 additional covariates in the model
making it very difficult to keep the model manageable regarding computational
costs but also interpretability. Therefore, only main effects are considered
for the distributional forest and the boosted GAMLSS which both automatically
select covariates while meteorological expert knowledge was applied to 
select covariates and specify interactions for the prespecified GAMLSS.
Therefore,

\subsubsection*{Tuning parameters}
Setting tuning parameters of a forest model can have a major influence
not only on the accuracy of the model but also on the computational 
complexity. 
For the application in Section~\ref{sec:precipitation} we have 
selected the values of these parameters (listed in Section~\ref{sec:evaluation}) 
based on advice from literature and our own experience.

\cite{Hastie+Tibshirani+Friedman:2001} and \cite{Breiman:2001} recommend to 
build full-grown trees. Therefore, low values should be chosen for the stopping 
criteria \code{minsplit}, \code{minbucket} and \code{alpha}.
However, in a distributional model it also has to be considered that a 
minimal number of observations is required to allow for a reasonable 
estimation of parameters in each node. For that reason the minimal number
of observations for which statistical tests are performed in the 
distributional forest in Section~\ref{sec:precipitation} is set to
50 (\code{minbucket}) and the minimal number of observations on which
parameters are estimated is set to 20 (\code{minsplit}).

As stated by \cite{Breiman:2001} and \cite{Hastie+Tibshirani+Friedman:2001} 
\hl{random forests do not overfit, not even for an increasing number of 
trees} (\code{ntree}). \hl{Hence, it is recommended to build forests consisting 
of a large number of trees} (\code{ntree}) \hl{as this generally leads to an increase 
in accuracy. However, this also comes along with increasing computational costs. 
Therefore, ``a good choice results from a trade-off between computational 
complexity and accuracy''} \citep{Biau+Scornet:2016}.
\hl{Following this advice we chose to build forests consisting of 100 trees. 
However, increasing this tuning parameter could improve the performance
of the distributional forest without the problem of overfitting.}

%% TO DOs:
%% East Tyrol
\subsubsection*{East Tyrol}
Figure~\ref{fig:all-map} shows that, contrary to the rest of Tyrol, 
for the majority of stations in East Tyrol the prespecified GAMLSS performs
best in this particular setting (learning data: 1985--2008, testing data:
2009--2012). East Tyrol is located south of the main Alpine Ridge and 
exhibits different climatological characteristics than the part north of 
the Alpine Ridge. It seems that the choice of covariates and their interactions 
we made for the prespecified GAMLSS method is particularly adequate for the 
region of East Tyrol for this specific analysis. However, the advantages of 
the prespecified GAMLSS models in this particular setting (training: 1985--2008, 
validation: 2009--2012) even out in a cross-validation framework for most of 
the stations in East Tyrol. These results are presented in the online supplement 
which comes along with our manuscript. There, more extensive evaluations
are carried out for 15 observation stations, including a cross-validation
as presented for station Axams in this paper.

\subsubsection*{Choice of station Axams}
%% choice of station Axams mit Verweis auf Supplement stationwise
Axams was chosen for the more extensive evaluations presented
in this paper as it is the observation station being closest 
to Innsbruck, the capital of Tyrol and the work place of 
three of the authors, but not based on any other particular 
features. Furthermore, it has been checked, whether it 
is an adequate representative of all stations regarding 
the results of our evaluations. This is now also 
confirmed by further analysis in the online supplement which
comes along with our manuscript.
There, the same evaluations as for station Axams have been carried out for 14 other
stations, including some of those where the forest does not perform better than
the other models in the setting as presented in Figure~\ref{fig:all-map}.  
It shows that the differences in performance even out in the cross-validation 
framework. There, the forest performs at least equally well compared to the 
best of the other models at most stations.

\subsubsection{Distributional specifications}
%% choice of distribution mit Verweis auf Supplement distributions
Moreover, as mentioned before we have tested robustness against 
distributional misspecification. The results show that applying a 
left-censored at zero logistic distribution or a Gaussian hurdle model 
instead of the left-censored at zero Gaussian distribution, as presented 
in the paper, does not lead to any significant improvements in goodness of 
fit of the applied models. The exact results are provided in an online 
supplement which comes with our manuscript.

\subsubsection*{Computational difficulties}
%% GAMLSS errors with references gamboostLSS Hofner et al.
As stated by \cite{Hofner+Mayr+Schmid:2016} the AIC-based variable selection 
methods implemented in the \textsf{R} package \textbf{gamlss} ``can be unstable, 
especially when it comes to selecting possibly different sets of variables for 
multiple distribution parameters.'' We have noticed computational problems when
applying the function \textbf{gamlss} in certain settings within the cross-validation
framework as it did not succeed in fitting a model. In this cases the method
was not taken into consideration in the comparison of all applied models.

>>>>

To assess the predictive performance, a full cross-validation is carried out
rather than relying on just the one fixed test set for the years 2009--2012.
\hl{For each setting in the considered 10 times 7-fold cross-validation the 
available 28~years are split into 7~subsets of 4~randomly selected and 
therefore mostly not subsequent years.}
The models are learned on 6~folds (= 24~years) and evaluated on the 7-th fold
(= 4~years) using the average CRPS across all observations.
The resulting 10~CRPS skill scores are displayed by boxplots in
Figure~\ref{fig:axams-crps} using EMOS as the reference model (horizontal
line at a CRPSS of 0). Both GAMLSS models and the distributional forest 
perform distinctly better than the EMOS model. While the two GAMLSS lead to an improvement 
of around 4~percent, the distributional forest has a slightly higher improvement
of around 5.5~percent in median.

@@

\subsubsection*{Variable Selection}
\hl{Generally, there are many possibilities how to specify the variables that are
to be included in a model. Especially for a low number of variables all types
of interactions can be employed and mixed effect models and separations into 
main effects and interaction effects can be considered allowing for easier 
interpretations} (see for example \cite{Wood+Scheipl+Faraway:2013}, 
\cite{Goicoa+Adin+Ugarte:2018}, \cite{Ugarte+Adin+Goicoa:2017}).
\hl{However, in the application presented in} Section~\ref{sec:precipitation}
\hl{alltogether 80 variables are available. Even if only paired interactions were
considered this would already lead to 3160 additional covariates in the model
making it very difficult to keep the model manageable regarding computational
costs but also interpretability. Therefore, only main effects are considered
for the distributional forest and the boosted GAMLSS which both automatically
select covariates while meteorological expert knowledge was applied to 
select covariates and specify interactions for the prespecified GAMLSS.}

\subsubsection*{Tuning parameters}
\hl{Setting tuning parameters of a forest model can have a major influence
not only on the accuracy of the model but also on the computational 
complexity. 
For the application in} Section~\ref{sec:precipitation} \hl{we have 
selected the values of these parameters} (listed in Section~\ref{sec:evaluation}) 
based on advice from literature and our own experience.

\cite{Hastie+Tibshirani+Friedman:2001} and \cite{Breiman:2001} \hl{recommend to 
build full-grown trees. Therefore, low values should be chosen for the stopping 
criteria} \code{minsplit}, \code{minbucket} and \code{alpha}.
\hl{However, in a distributional model it also has to be considered that a 
minimal number of observations is required to allow for a reasonable 
estimation of parameters in each node. For that reason the minimal number
of observations for which statistical tests are performed in the 
distributional forest in} Section~\ref{sec:precipitation} is set to
50 (\code{minbucket}) \hl{and the minimal number of observations on which
parameters are estimated is set to 20} (\code{minsplit}).

As stated by \cite{Breiman:2001} and \cite{Hastie+Tibshirani+Friedman:2001} 
\hl{random forests do not overfit, not even for an increasing number of 
trees} (\code{ntree}). \hl{Hence, it is recommended to build forests consisting 
of a large number of trees as this generally leads to an increase 
in accuracy. However, it also comes along with increasing computational costs. 
Therefore, ``a good choice results from a trade-off between computational 
complexity and accuracy''} \citep{Biau+Scornet:2016}.
\hl{Following this advice we chose to build forests consisting of 100 trees.}

\subsubsection*{East Tyrol}
Figure~\ref{fig:all-map} \hl{shows that, contrary to the rest of Tyrol, 
for the majority of stations in East Tyrol the prespecified GAMLSS performs
best in this particular setting (learning data: 1985--2008, testing data:
2009--2012). East Tyrol is located south of the main Alpine Ridge and 
exhibits different climatological characteristics than the part north of 
the Alpine Ridge. It seems that the choice of covariates and their interactions 
we made for the prespecified GAMLSS method is particularly adequate for the 
region of East Tyrol for this specific analysis. However, the advantages of 
the prespecified GAMLSS models in this particular setting (training: 1985--2008, 
validation: 2009--2012) even out in a cross-validation framework for most of 
the stations in East Tyrol. These results are presented in Supplement 2 
(Stationwise Evaluation). There, the exact same evaluations as in} 
Section~\ref{sec:axams} \hl{for station Axams have been carried out 
for 14 other stations.}

\subsubsection*{Choice of station Axams}
%% choice of station Axams mit Verweis auf Supplement stationwise
\hl{Axams was chosen for the more extensive evaluations presented
in this paper as it is the observation station being closest 
to Innsbruck, the capital of Tyrol and the work place of 
three of the authors, but not based on any other particular 
features. Furthermore, it has been checked, whether it 
is an adequate representative of all stations regarding 
the results of our evaluations. This is also confirmed by 
the extended analysis in Supplement 2 (Stationwise Evaluation)
where the 15 considered stations also include some of those where 
the forest does not perform better than the other models in the setting 
presented in} Figure~\ref{fig:all-map}. \hl{It shows that the differences in 
performance even out in the cross-validation framework where the forest 
performs at least equally well compared to the best of the other models 
at most stations.}

\subsubsection{Distributional specifications}
%% choice of distribution mit Verweis auf Supplement distributions
\hl{Choosing an adequate distributional family is a crucial step and forms
the basis required to build a well fitting model. Therefore, this
decision should be made thoughtfuly. 
In the presented application we applied a Gaussian left-censored at
zero distribution as this has proven to be a well working choice in
earlier literature} \citep{Stauffer+Mayr+Messner:2017}. 
\hl{But to test for robustness against distributional misspecifation 
different distributional assumptions have been considered as well. 
In Supplement 1 (Different Response Distributions) the exact
same evaluations as in} Section~\ref{sec:all} \hl{are carried out for 15
observation stations, once employing a Gaussian hurdle model and once
with a left-censored at zero logistic distribution in order to account
for heavier tails. Under both assumptions the obtained results 
are similar to those presented in the paper applying a left-censored
at zero Gaussian distribution. However, comparing the two censored distributions,
at many stations the models perform not as well under the logisctic distribution 
as under the Gaussian distribution. The hurdle model leads to slight improvements 
in some settings, but this comes with more variablity (especially for EMOS and the 
boosted GAMLSS).
Alltogether, the distributional forest has proven to be very robust against
distributional misspecifications.}

\subsubsection*{Computational difficulties}
%% GAMLSS errors with references gamboostLSS Hofner et al.
As stated by \cite{Hofner+Mayr+Schmid:2016} \hl{the AIC-based variable selection 
methods implemented in the} \textsf{R} package \textbf{gamlss} \hl{``can be unstable, 
especially when it comes to selecting possibly different sets of variables for 
multiple distribution parameters.'' We have noticed computational problems when
applying the function} \textbf{gamlss} \hl{in certain settings within the cross-validation
framework as it did not succeed in fitting a model. In this cases the method
was not taken into consideration in the comparison of all applied models.}
